Tidy data structure to support exploration and modeling of temporal-context data

### 2018-08-19 main.Rmd
done with sec 1.2

A collection of time series can often be arranged along multivariate data of nested and crossed factors, which is also known as hierarchical and grouped time series [@fpp]. For example, from an operational point of view, a manufacturer can add up every store’s sales by region, by state, and by country along a geographical tree; alternatively, one can group the sales for each product together based on common attributes such as store, brand, price range and so forth, resulting in a non-hierarchical structure. The former gives an example of hierarchical time series, and the latter shows grouped time series. Hierarchical time series can also be derived using hierarchical clustering.

The R package **hts** [@R-hts] is the implementation of this type of time series. But it is frustrating to create and work with such data objects due to non-extensible and non-modified. Variational data aggregation (some groups no need aggregation). [R]

@@

The "key" identifies units or subjects that are observed over time in a data table, which is typically a priori known structure by users. The absence of key only occurs to a univariate case where the key is implicit. Multiple units must associate with an explicit "key" to allow for identification. The key is not constrained to a single column, but is comprised of multiple columns. When involving multiple variables, one need to distinguish nesting and crossing data variables. "Nesting" refers to a scenario where a variable is nested within another when each category of the former variable co-occurs with only one category of the latter (for example, `country` nested within `continent`); on the other hand, "crossing" means that a variable is crossed with another when every category of one variable co-occurs with every category of the other (for example, `country` crosses with `gender`). Nesting is a special case of crossing. The former exhibits a hierarchical ordering with lower-level category (children) nested within higher-level category (parent) (more categories in lower level); whereas the ordering of the latter makes no difference. Whilst constructing tidy temporal data, users also need to specify the key that identifies the structure on the data. Rather than creating a separate object that contains the data structure, like what the **hts** does, we propose a syntax-based interface to directly deal with variables, so that it takes advantage of data semantics. The expression of a vertical bar (`|`) (can be interpreted probabilistically as the conditional distribution for its parent) is used to indicate nesting variables, while a comma (`,`) for crossing variables. @wilkinson2006grammar discussed the distinction between nesting and crossing with respect to facets in graphics, and he used `/` and `*` instead for nested and crossed expressions. A crossing structure displays all unique and possible combinations of crossed variables, including those not found in the data, in a graphical layout. However, tidy temporal data finds only the combinations that occur in the data and intentionally respects the structure of incomplete combinations.

>>>>

A collection of time series can often be intrinsically arranged along multivariate data of nested and crossed factors, which is also known as hierarchical and grouped time series [@fpp]. For example, from an operational point of view, a manufacturer can add up every store’s sales by region, by state, and by country along a geographical tree; alternatively, one can group the sales for each product together based on common attributes such as store, brand, price range and so forth, resulting in a non-hierarchical structure. The former gives an example of hierarchical time series, and the latter shows grouped time series. And even when collections of time series are absent from hierarchical structure, it is possible to arrange them empirically into a hierarchy such as regression tree [@sankaran_interactive_2017].

The R package **hts** [@R-hts] is the implementation of structured collections of time series, and provides tools for visualizing, analyzing, and forecasting such data. But it is a pain to configure node list or grouping matrix as the increase in hierarchic depth and grouping factors. This is resulted from the fact that the structure specification is separated from data itself and represented numerically, which is rather counterintuitive from the user's perspective.

@@

The "key" identifies units or subjects that are observed over time in a data table, which is typically a priori known structure by users. The absence of key only occurs to a univariate case where the key is implicit. Multiple units must associate with an explicit "key" to allow for identification. The key is not constrained to a single column, but is comprised of multiple columns. When involving multiple variables, one need to distinguish nesting and crossing data variables. "Nesting" refers to a scenario where a variable is nested within another when each category of the former variable co-occurs with only one category of the latter (for example, `country` nested within `continent`); on the other hand, "crossing" means that a variable is crossed with another when every category of one variable co-occurs with every category of the other (for example, `country` crosses with `gender`). Nesting is a special case of crossing. The former exhibits a hierarchical ordering with lower-level category (children) nested within higher-level category (parent) (more categories in lower level); whereas the ordering of the latter makes no difference. Whilst constructing tidy temporal data, users also need to specify the key that identifies the structure on the data. Rather than creating a separate object that contains the data structure, like what the **hts** does, we propose a syntax-based interface to directly deal with variables, so that it takes advantage of data semantics. The expression of a vertical bar (`|`) (can be interpreted probabilistically as the conditional distribution for its parent) is used to indicate nesting variables (reflect depths), while a comma (`,`) for crossing variables. @wilkinson2006grammar discussed the distinction between nesting and crossing with respect to facets in graphics, and he used `/` and `*` instead for nested and crossed expressions. A crossing structure displays all unique and possible combinations of crossed variables, including those not found in the data, in a graphical layout. However, tidy temporal data finds only the combinations that occur in the data and intentionally respects the structure of incomplete combinations.
